---
title: "Summarizing covid-19 situation from 1.10.20 to 31.01.21"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

#### Clean dataset

Read data we are interested in, then those that are relative to the following couple of parameters:

-   regione: Sicilia

-   data: 1st October 2020 to 1st February 2021

```{r df, echo=FALSE, message=FALSE, warning=FALSE}
##################
#### Project SMSD
##################

library(ggplot2)


df_global <- data.frame(read.csv("https://raw.githubusercontent.com/pcm-dpc/
COVID-19/master/dati-regioni/dpc-covid19-ita-regioni.csv"))

#select data about Sicily and the period to be considered
df_sicily_secondwave <- df_global[which(df_global$denominazione_regione == "Sicilia" & 
                                          (df_global$data >= "2020-09-24T10:00:00" & df_global$data <= "2021-02-15T10:00:00")), ]

# adjust number of swabs
df_sicily_secondwave[which(df_sicily_secondwave$data >= "2021-01-15" & df_sicily_secondwave$data <= "2021-02-15"), ]$tamponi <- 
          df_sicily_secondwave[which(df_sicily_secondwave$data >= "2021-01-15" & df_sicily_secondwave$data <= "2021-02-15"), ]$tamponi -
          df_sicily_secondwave[which(df_sicily_secondwave$data >= "2021-01-15" & df_sicily_secondwave$data <= "2021-02-15"),]$tamponi_test_antigenico_rapido

# remove unused columns
df_extended <- df_sicily_secondwave[ , -which(names(df_sicily_secondwave) %in% c("stato", "codice_regione", "denominazione_regione", "lat", "long", "note"))]
df_extended <- df_extended[ , -c(17:24)]

# put data in Date format
df_extended$data <- as.Date(df_extended$data,  "%Y-%m-%d")


# add columns accounting for the number of daily swabs, daily deaths,
# people daily discharged from the hospital and the number of ICU of one week before 
df_extended$nuovi_tamponi <-NA
df_extended$nuovi_decessi <-NA
df_extended$nuovi_dimessi <- NA
df_extended$terapia_intensiva_prev <- NA
df_extended$ricoverati_con_sintomi_prev <- NA


for(x in 8:nrow(df_extended)) {
  df_extended$nuovi_tamponi[x] <- df_extended$tamponi[x] - df_extended$tamponi[x-1] # the number of new swabs carried out
  df_extended$nuovi_decessi[x] <- df_extended$deceduti[x] - df_extended$deceduti[x-1] # daily deaths
  df_extended$nuovi_dimessi[x] <- df_extended$dimessi_guariti[x] - df_extended$dimessi_guariti[x-1] # variation in the number of people discharged from the hospital
  df_extended$terapia_intensiva_prev[x] <- df_sicily_secondwave$terapia_intensiva[x-7]
  df_extended$ricoverati_con_sintomi_prev[x] <- df_sicily_secondwave$ricoverati_con_sintomi[x-7]
}

# remove first rows 
df_extended <- df_extended[-c(1:7),]

# index reordering
row.names(df_extended) <- NULL 

# add color
df_extended$color <-NA
df_extended$color[1:36]<- "bianco"
df_extended$color[c(37:59, 89:91, 101:108, 124:137)]<- "arancione"
df_extended$color[c(60:84,99,100)]<- "giallo"
df_extended$color[c(85:88,92:98, 109:123)]<- "rosso"

df <- df_extended[1:123,]
df$data <- as.Date(df$data,  "%Y-%m-%d")


library(ggplot2)
library(patchwork)


newpos_col1 <- ggplot() +
          geom_line(data = df, aes(x = data, y = nuovi_positivi, group = 1)) +
          geom_point(data = df, aes(x = data, y = nuovi_positivi, group = 1),col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))])+ 
          labs(x = "Time", y = "New positives") +
          scale_x_date(date_breaks = '5 days',
                       date_labels = '%Y-%m-%d') +
          theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

newpos_col1

# adjust color
df_extended$color[c(99,100)]<- "arancione"
df_extended$color[c(89:91)]<- "rosso"


```
1) Some useful analysis, correlation and plots about covid data from 1/10/20 to 31/01/21 


```{r}
    plotPositiveDeath <- ggplot() +
    geom_line(data = df, aes(x = data, y = (nuovi_decessi/nuovi_positivi)*100, color="rapporto percentuale tra deceduti e positivi", group = 1)) +
    labs(x = "Time", y = "% ratio of positive to deceased", title = "Fatality rate") +
    scale_x_date(date_breaks = '5 days',
                 date_labels = '%Y-%m-%d') +
    #scale_y_continuous(limits = c(0,100))+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "none", plot.title = element_text(hjust = 0.5))
  
  plotPositiveDeath
```





```{r}

  plotPositiveSwabs <- ggplot() +
    geom_line(data = df, aes(x = data, y = (nuovi_positivi/nuovi_tamponi)*100, color="rapporto percentuale tra deceduti e positivi", group = 1)) +
    labs(title = "Positivity rate", x = "Time", y = "% ratio of tested swabs to positives") +
    scale_x_date(date_breaks = '5 days',
                 date_labels = '%Y-%m-%d') +
    #scale_y_continuous(limits = c(0,100))+ 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "none", plot.title = element_text(hjust = 0.5))
  plotPositiveSwabs
```



```{r}
    theme_set(theme_bw())
    plotIntensive <- ggplot() +
    geom_line(data = df, aes(x = data, y = terapia_intensiva, color="Places occupied in intensive care", group = 1)) +
    geom_line(data = df, aes(x = data, y = ricoverati_con_sintomi, color="hospitalised with symptoms", group = 1)) +
    labs(title = "Hospitalised and places occupied in intensive terapy", x = "Time", y = "Number of places occupied") +
    scale_x_date(date_breaks = '5 days',
                 date_labels = '%Y-%m-%d') +
    #scale_y_continuous(limits = c(0,100))+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), plot.title = element_text(hjust = 0.5))
  plotIntensive
```


```{r}
plotSummary <- ggplot() +
    geom_line(data = df, aes(x = data, y = nuovi_positivi, color="Positives", group = 1))+
    geom_line(data = df, aes(x = data, y = nuovi_decessi, color="Deaths", group = 1))+
    geom_line(data = df, aes(x = data, y = terapia_intensiva, color="Intensive care", group = 1))+
    geom_line(data = df, aes(x = data, y = ricoverati_con_sintomi, color="Hospitalised with symptoms", group = 1)) +
    labs(title = "Summary of covid situation between 1/10/20 and 31/02/21", x = "Time", y = "Cases") +
    scale_x_date(date_breaks = '5 days',
                 date_labels = '%Y-%m-%d') +
    #scale_y_continuous(trans = "log10")+
    #scale_y_continuous(limits = c(0,100))+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), plot.title = element_text(hjust = 0.5))

plotSummary



```




```{r}
plotOtherSummary <- ggplot() +
    geom_line(data = df, aes(x = data, y = totale_positivi, color="Positives", group = 1))+
    geom_line(data = df, aes(x = data, y = deceduti, color="Deaths", group = 1))+
    geom_line(data = df, aes(x = data, y = terapia_intensiva, color="Intensive care", group = 1))+
    geom_line(data = df, aes(x = data, y = ricoverati_con_sintomi, color="Hospitalised with symptoms", group = 1)) +
    labs(title = "Summary of covid situation between 1/10/20 and 31/02/21", x = "Time", y = "Cases") +
    scale_x_date(date_breaks = '5 days',
                 date_labels = '%Y-%m-%d') +
    scale_y_continuous(trans = "log10")+
    #scale_y_continuous(limits = c(0,100))+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), plot.title = element_text(hjust = 0.5))
plotOtherSummary
```

Here we're trying to predict nuovi_positivi with LOESS model. LOESS combines much of the simplicity of linear least squares regression with the flexibility of nonlinear regression.

```{r}

model_0 <- lm(nuovi_positivi ~ nuovi_decessi + color, data = df)

#(model_0)

model_1 <- loess(nuovi_positivi ~ nuovi_decessi + isolamento_domiciliare + ricoverati_con_sintomi_prev, data = df,control = loess.control(surface = "direct"), nsigma = 1.96, span = 0.7)

model_loess <- loess.sd(y = df$nuovi_positivi, x = df$nuovi_decessi + df$ricoverati_con_sintomi_prev + df$isolamento_domiciliare)

?loess
smoothed_1 <- predict(model_1)
smoothed_2 <- predict(model_1, newdata = df_extended, type = "response")

smoothed_2 <- forecast(model_1, 15)
#df$nuovi_positivi
?forecast
ggplot() + 
  geom_line(data = df, aes(x = data, y = model_loess$x, color="Model", group = 1)) +
  geom_line(data = df, aes(x = data, y = model_loess$upper, color="upper", group = 1)) +
  geom_line(data = df, aes(x = data, y = model_loess$lower, color="lower", group = 1)) +
  geom_point(data = df, aes(x = data, y = nuovi_positivi, color="Real", group = 1),col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))]) +
  geom_line(data = df_extended, aes(x = data, y = nuovi_positivi, color="Real", group = 1)) +
  #geom_point(data = df_extended, aes(x = data, y = smoothed_2, color="ModelPredicted", group = 1)) +
  geom_line(data = df_extended, aes(x = data, y = smoothed_2, color="ModelPredicted", group = 1)) +
  geom_smooth(se = FALSE)+
  labs(title = "Model_V1", x = "Time", y = "Cases") +
    scale_x_date(date_breaks = '5 days',
                 date_labels = '%Y-%m-%d') +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), plot.title = element_text(hjust = 0.5))

ggplot() + 
  geom_point(data = df, aes(x = data, y = nuovi_positivi, color="Real", group = 1),col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))]) +
  geom_smooth(data = df, aes(x = data, y = df$nuovi_positivi))

#plot(smoothed_1)
```


Some model based on time series
```{r}
library(tidyverse)      # data manipulation and visualization
library(lubridate)      # easily work with dates and times
library(fpp2)           # working with time series data
library(zoo)
library(forecast)
library(HSAR)


#model_2 <- ts(nuovi_positivi ~ nuovi_decessi  + isolamento_domiciliare, data = df)



model_ts <- ts(data = df$nuovi_positivi, start =  1, end = 123)

model_ts2 <- ts(data = df$nuovi_positivi, start =  1, frequency = 123)

plot(model_ts2)
plot(model_ts)
fit <- auto.arima(model_ts)

plot(forecast(fit, 15))
a <- predict(fit, n.ahead=15)
plot(x = 1:15,y = a$pred)

plot(fit)

ggplot() + 
  #geom_line(data = df_extende, aes(x = data, y = nuovi_positivi, color="Real", group = 1))+
  geom_line(data = df_extended, aes(x = data, y = forecast(fit,14), color="Model", group = 1))

df_extended$data
forecast(fit,14)
plot(auto.arima(model_ts))

model_ts1 <- ts(cbind(df$nuovi_positivi, df$nuovi_decessi, df$isolamento_domiciliare), start =  1, end = 123)

model_ts1
plot(model_ts1)

fit <- sma(model_ts1)

?arima

```

TIME SERIES

```{r}
library(tidyverse)      # data manipulation and visualization
library(lubridate)      # easily work with dates and times
library(fpp2)           # working with time series data
library(zoo)
library(forecast)
library(HSAR)

df_sicily_global <- df_global[which(df_global$denominazione_regione == "Sicilia" & (df_global$data >= "2020-02-24T10:00:00" & df_global$data <= "2021-02-15T10:00:00")), ]
model_ts <- ts(data = df_sicily_global$nuovi_positivi, start =  1, end = 178, frequency = 2)
dec <- decompose(model_ts, type = "multiplicative")
names(dec)
fit <- nnetar(model_ts)
plot(dec)
plot(forecast(fit), 60)
autoplot(forecast(auto.arima(model_ts, d = 2), h = 20))
?auto.arima
model_ts3 <- ts(df$nuovi_positivi, start = c(10), end = c())

df_sicily_global$nuovi_positivi
df_sicily_global

```

