---
title: "Covid-19 - SMSD"
output:
  html_document: default
---
  
#### Goal
  
Build up a useful model in order to predict response variable Y starting from data provided.

-   response variable: number of new positives
-   data: Covid-19 spreading outbreak from the official website of Protezione Civile

#### Dataset

Load dataset.

```{r dataset, message=FALSE, echo = FALSE}
df_extended <- data.frame(read.csv("sicily_secondwave_covid.csv"))
df_extended$data <- as.Date(df_extended$data,  "%Y-%m-%d")

df <- df_extended[which(df_extended$data >= "2020-10-01" & df_extended$data <= "2021-01-31"),]
df$data <- as.Date(df$data,  "%Y-%m-%d")

```

#### Clean dataset

Read data we are interested in, then those that are relative to the following couple of parameters:
  
-   regione: Sicilia

-   data: 1st October 2020 to 1st February 2021

```{r prior analysis, echo=FALSE, message=FALSE, warning=FALSE}
library("ggcorrplot")
library("corrplot")
cols <- c("ricoverati_con_sintomi","terapia_intensiva","totale_ospedalizzati", "isolamento_domiciliare", "totale_positivi", "dimessi_guariti", "deceduti", "tamponi","nuovi_positivi")
M=cor(df_extended[,cols])
colnames(M) <- c("A", "B", "C", "D", "E", "F", "G", "H", "I")
rownames(M) <- paste0(colnames(M), ". ", gsub("_", " ", cols))

col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))

png("images/variables_corrplot.png", width=10,height=8,units="in",res=1200)
corrplot(M,cl.lim=c(0,1),is.corr=FALSE, method="color",col=col(200), 
     diag=FALSE, type="upper", addCoef.col = "black", tl.col = "black",tl.srt = 360, tl.offset = 1, tl.cex=1.1)
dev.off()

corrplot(M,cl.lim=c(0,1),is.corr=FALSE, method="color",col=col(200), 
     diag=FALSE, type="upper", addCoef.col = "black", tl.col = "black",tl.srt = 360, tl.offset = 1, tl.cex=1.1)

```

```{r plot response var against time, echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(patchwork)

ggplot() +
  geom_line(data = df, aes(x = data, y = nuovi_positivi, group = 1)) +
  geom_point(data = df, aes(x = data, y = nuovi_positivi, group = 1),col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))], size=2)+ 
  labs(x = "Time", y = "New positives") +
  scale_x_date(date_breaks = '5 days',
               date_labels = '%Y-%m-%d') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


ggsave(paste0( "images/response_variable.png"), width = 20, height = 8, dpi = 150)


```

#### Perform some explanatory analysis for your data.

- Observe correlation plot between variables

```{r corrplot, echo=FALSE, message=FALSE, warning=FALSE}
library("ggcorrplot")
library("corrplot")
cols <- c("ricoverati_con_sintomi","nuovi_decessi", "terapia_intensiva","nuovi_tamponi_pcr", "nuovi_positivi")
M=cor(df[,cols])
colnames(M) <- c("A", "B", "C", "D", "E")
rownames(M) <- paste0(colnames(M), ". ", gsub("_", " ", cols))
corrplot(M, method="number",tl.col="#a13c28", tl.srt = 360, tl.offset = 1, tl.cex=1.1)

```

- See how response variable changes wrt explanatory variables provided.

```{r variables, echo=FALSE, message=FALSE, warning = FALSE, comment = FALSE}
png("images/corr_scatterplot_response_variable.png", width=8.25,height=5.25,units="in",res=1200)
par(mfrow=c(2,3))
plot(nuovi_positivi ~ ricoverati_con_sintomi, data = df, pch =16, xlab = "Hospitalized with symptoms", ylab = "New positives",col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))])
plot(nuovi_positivi ~ terapia_intensiva, data = df, pch =16, xlab = "ICU", ylab = "New positives", col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))]) # ICU stands for the number of patients in intensive care
plot(nuovi_positivi ~ isolamento_domiciliare, data = df, pch =16, xlab = "Home isolation", ylab = "New positives",col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))])
plot(nuovi_positivi ~ nuovi_decessi, data = df, pch =16, xlab = "Daily number of deaths", ylab = "New positives",col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))])
plot(nuovi_positivi ~ nuovi_dimessi, data = df, pch =16, xlab = "Daily number of discharged", ylab = "New positives", col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))])
plot(nuovi_positivi ~ nuovi_tamponi_pcr, data = df, pch =16, xlab = "PCR test", ylab = "New positives",col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))])

dev.off()

png("images/corr_scatterplot_response_variable_lag.png", width=8.25,height=5.25,units="in",res=1200)
par(mfrow=c(2,2))
plot(nuovi_positivi ~ ricoverati_con_sintomi_prev, data = df, pch =16, xlab = "Hospitalized with symptoms (15d lag)", ylab = "New positives",col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))])
plot(nuovi_positivi ~ terapia_intensiva_prev, data = df, pch =16, xlab = "ICU (15d lag)", ylab = "New positives", col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))]) # ICU stands for the number of patients in intensive care

plot(nuovi_positivi ~ nuovi_tamponi_pcr_prev, data = df, pch =16, xlab = "PCR test (15d lag)", ylab = "New positives",col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))])
plot(nuovi_positivi ~ nuovi_positivi_prev, data = df, pch =16, xlab = "New positives  (15d lag)", ylab = "New positives",col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))])

dev.off()

```

-   See how response variable changes wrt covariates considered.

```{r boxplot_scatterplot, echo =FALSE}
library(ggplot2)
ggplot(df,aes(x=factor(color),y=nuovi_positivi))+
  geom_boxplot(outlier.color="black", fill=c("#fc6b03","white","#f2d729","#b3190b"))+
  labs(x = " ", y = "Daily number of new positives")

plot(df$data, df$nuovi_positivi,col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))], pch=16, xlab="Time", ylab="New positives")
legend("topleft",legend=c( "Zona arancione","Zona bianca", "Zona gialla", "Zona rossa"),
       col=c("#fc6b03","#cfcaca","#f2d729","#b3190b"),pch=16, bty="n",cex=0.8,pt.cex=1.2) 
```





#### Quality of the data

Check for missing values in the data frame

```{r na, echo=FALSE}
ncols <- dim(df)[2]
na_count <- sapply(df[,-c((ncols-15):ncols)], function(y) sum(length(which(is.na(y)))))

na_count <- data.frame( gsub("_", " ", names(na_count)),na_count)
colnames(na_count) <- c("variable", "missing values")

library(gridExtra)
library(grid)
color_df <- data.frame(color = c("#bed0e8", "#809cc2"), stringsAsFactors = FALSE)
my_table_theme <- ttheme_default(core=list(bg_params = list(fill = color_df$color[1:2], col=NA)), colhead =list(bg_params=list(fill ="#5680ba")))

png("images/missing_values_table.png", width=4.5,height=6.5,bg = "white",units="in",res=1200)
grid.table(na_count, rows=NULL)
dev.off()


```

-   The most important missing value is about the admission to the ICU and for this reason that value has been removed and it has been considered the number of ICU per day.

-   Discussion about the color

-   Write something about the number of swabs considered (?)

### Correlations
-   Understand how each potential predictor relates to the outcome but also to every other predictor.

We observe the following variables: 
-   number of hospedalized with symptoms 
-   number of ICU 
-   number of deaths 
-   number of swabs

```{r ricoverati, echo=FALSE}
par(mfrow=c(2,3))

plot(ricoverati_con_sintomi ~ terapia_intensiva, data = df, pch =16, xlab = "ICU", ylab = "Hospitalized with symptoms") # ICU stands for the number of patients in intensive care

plot(ricoverati_con_sintomi ~ nuovi_decessi, data = df, pch =16, xlab = "Daily number of deaths", ylab = "Hospitalized with symptoms")

plot(ricoverati_con_sintomi ~ nuovi_tamponi_pcr, data = df, pch =16, xlab = "Number of PCR test", ylab = "Hospitalized with symptoms")


plot(terapia_intensiva ~ nuovi_decessi, data = df, pch =16, xlab = "Daily number of deaths", ylab = "ICU")

plot(terapia_intensiva ~ nuovi_tamponi_pcr, data = df, pch =16, xlab = "Number of PCR test", ylab = "ICU")

plot(nuovi_decessi ~ nuovi_tamponi_pcr, data = df, pch =16, xlab = "Number of PCR test", ylab = "Daily number of deaths")

```

<br>
  
  
```{r swabs_positives, echo=FALSE}
ggplot() +
  geom_line(data = df, aes(x = data, y = log(nuovi_positivi), color="new positives", group = 1),) +
  geom_point(data = df, aes(x = data, y = log(nuovi_positivi), group = 1))+ 
  geom_line(data = df, aes(x = data, y = log(nuovi_tamponi_pcr), color="PCR test", group = 1)) +
  geom_point(data = df, aes(x = data, y = log(nuovi_tamponi_pcr),color="PCR test", group = 1)) +
  geom_line(data = df, aes(x = data, y = log(nuovi_decessi), color="deaths", group = 1)) +
  geom_point(data = df, aes(x = data, y = log(nuovi_decessi),color="deaths", group = 1)) +
  geom_line(data = df, aes(x = data, y = log(terapia_intensiva), color="ICU", group = 1)) +
  geom_point(data = df, aes(x = data, y = log(terapia_intensiva),color="ICU", group = 1)) +
  geom_line(data = df, aes(x = data, y = log(isolamento_domiciliare), color="home isolation", group = 1)) +
  geom_point(data = df, aes(x = data, y = log(isolamento_domiciliare),color="home isolation", group = 1)) +
  scale_x_date(date_breaks = '5 days', date_labels = '%Y-%m-%d') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_colour_manual(values = c("new positives" = "black", "PCR test" ="#2f6d80", "deaths" = "#490e96", "ICU" = "#1a8f2f", "home isolation" = "#851128")) 


```

<br>
  
```{r correlation, echo=FALSE, message=FALSE, warning=FALSE}
library("ggcorrplot")
library("corrplot")
cols <- c("ricoverati_con_sintomi","nuovi_decessi", "terapia_intensiva","nuovi_tamponi_pcr", "nuovi_positivi")
M=cor(df[,cols])
colnames(M) <- c("A", "B", "C", "D", "E")
rownames(M) <- paste0(colnames(M), ". ", gsub("_", " ", cols))
corrplot(M, method="number",tl.col="#a13c28", tl.srt = 360, tl.offset = 1, tl.cex=1.1)

```



## Building models

Then we start exploring the previous correlation, since the number of swabs carried out is one of the main factor which determines the number of new positives. Indeed, the number of positives is strictly contained in the interval
[0, n. of swabs].

```{r mod1, echo =FALSE, message=FALSE, warning=FALSE}
# first model
mod1 <- lm(nuovi_positivi ~ nuovi_tamponi_pcr, data = df)
summary(mod1)

par(mfrow=c(1,2))
plot(nuovi_positivi ~ nuovi_tamponi_pcr, data = df, pch =16, xlab = "Number of PCR test", ylab = "Number of new positives",col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))])
abline(coef(mod1), col=3, lwd=2)

## Model checking - Residual plots
pred <- predict(mod1)
y_res <- df$nuovi_positivi - pred
plot(pred, y_res, pch=16, col=4, xlab ="predictions", ylab = "residuals")

```

We can observe that the residuals tend to grow for increasing values of our response variable, which corresponds to the days in which there are some restrictions.
A second model which accounts for the presence or absence of restrictions:
  
  
```{r mod2, echo =FALSE, message=FALSE, warning=FALSE}
mod2 <- lm(nuovi_positivi ~ nuovi_tamponi_pcr + color, data = df)
summary(mod2)

par(mfrow=c(1,2))
plot(nuovi_positivi ~ nuovi_tamponi_pcr, data = df, pch =16, xlab = "Number of PCR test", ylab = "Number of new positives",col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))])

points(df$nuovi_tamponi_pcr, predict(mod2), data = df,pch=20, xlab = "Number of PCR test", ylab = "Number of new positives")

pred <- predict(mod2)
y_res <- df$nuovi_positivi - pred
plot(pred, y_res, pch=16, col=4, xlab ="predictions", ylab = "residuals")

```

Now we order and see how the residuals change wrt to time, or the date we are considering:
  
  ```{r mod2 errors, echo =FALSE, message=FALSE, warning=FALSE}
## Model checking - Residual plots
library(modelr)
library(ggplot2)
df <- df %>% 
  add_residuals(mod2)

df %>% 
  ggplot(aes(data, resid)) + 
  geom_ref_line(h = 0) + 
  geom_line(col=1) + 
  geom_smooth(se = FALSE, span = 0.20, col=3)

```


It seems there is still a kind of pattern in the residuals, initially we underestimate the number of new positives, then in November the residuals increase and they increase again in January.

The previous models try to underline the relevance of the region color, which was the variable less correlated to the number of new positives in the previous correlation plot.

## Building models

We start building our model looking at the relationship between our response variable and its highest correlated predictor (which is given by people hospitalized with symptoms, maybe because the highest the number of people hospitalized, the higher the contagiousness and as a consequence the higher the number of new positives).

### Linear models

```{r linear models, echo=FALSE, message=FALSE, warning=FALSE}
# Try some linear models to see which are the variables which significantly explain the variability of y
lm_mod1 <- glm(nuovi_positivi ~ ricoverati_con_sintomi, data = df, family="gaussian")
lm_mod2 <- glm(nuovi_positivi ~ ricoverati_con_sintomi + color, data = df, family="gaussian")
lm_mod3 <- glm(nuovi_positivi ~ ricoverati_con_sintomi * color, data = df, family="gaussian")
lm_mod4 <- glm(nuovi_positivi ~ ricoverati_con_sintomi + color + data , data = df, family="gaussian")

summary(lm_mod1)
summary(lm_mod2)
summary(lm_mod3)
summary(lm_mod4)

ggplot() +
  geom_point(data = df, aes(x = data, y = nuovi_positivi, color="nuovi positivi", group = 1),col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))])+ 
  geom_line(data = df, aes(x = data, y = predict(lm_mod1), color = "model 1", group = 1)) +
  geom_line(data = df, aes(x = data, y = predict(lm_mod2), color = "model 2", group = 1)) +
  geom_line(data = df, aes(x = data, y = predict(lm_mod3), color = "model 3", group = 1)) +
  geom_line(data = df, aes(x = data, y = predict(lm_mod4), color = "model 4", group = 1)) +
  
  labs(x = "Time", y = "New positives") +
  scale_x_date(date_breaks = '5 days',
               date_labels = '%Y-%m-%d') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


AIC(lm_mod1,lm_mod2,lm_mod3,lm_mod4)
```
### Generalized linear models

```{r glm, echo=FALSE, message=FALSE, warning=FALSE}

glm_mod1 <- glm(nuovi_positivi ~ ricoverati_con_sintomi + color, data = df, family=poisson(link = "log"))
glm_mod2 <- glm(nuovi_positivi ~ ricoverati_con_sintomi + color + data, data = df, family=poisson(link = "log"))
glm_mod3 <- glm(nuovi_positivi ~ ricoverati_con_sintomi * color + data, data = df, family=poisson(link = "log"))
glm_mod4 <- glm(nuovi_positivi ~ (ricoverati_con_sintomi + terapia_intensiva) * color + data, data = df, family=poisson(link = "log"))

summary(glm_mod1)
summary(glm_mod2)
summary(glm_mod3)
summary(glm_mod4)

ggplot() +
  geom_point(data = df, aes(x = data, y = nuovi_positivi, color="nuovi positivi", group = 1),col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))])+ 
  geom_line(data = df, aes(x = data, y = predict(glm_mod1,type="response"), color = "model 1", group = 1)) +
  geom_line(data = df, aes(x = data, y = predict(glm_mod2,type="response"), color = "model 2", group = 1)) +
  geom_line(data = df, aes(x = data, y = predict(glm_mod3,type="response"), color = "model 3", group = 1)) +
  geom_line(data = df, aes(x = data, y = predict(glm_mod4,type="response"), color = "model 4", group = 1)) +
  
  labs(x = "Time", y = "New positives") +
  scale_x_date(date_breaks = '5 days',
               date_labels = '%Y-%m-%d') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

AIC(glm_mod1, glm_mod2, glm_mod3, glm_mod4)


```

```{r glm residuals, echo =FALSE, message=FALSE, warning=FALSE}
## Model checking - Residual plots
library(modelr)
library(ggplot2)

df$resid <- df$nuovi_positivi -  predict(glm_mod4,type="response")

ggplot(df, aes(data, resid, colour = color)) + 
  geom_ref_line(h = 0) + 
  geom_line()

df %>% 
  ggplot(aes(data, resid)) + 
  geom_ref_line(h = 0) + 
  geom_line(col=1) + 
  geom_smooth(se = FALSE, span = 0.20, col=3)

```


Now we add the another covariate which accounts for the number of ICU of one week before
```{r glm prev ICU, echo=FALSE, message=FALSE, warning=FALSE}

glm_mod5 <- glm(nuovi_positivi ~ ricoverati_con_sintomi * color + data + ricoverati_con_sintomi_prev + nuovi_tamponi_pcr, data = df, family=poisson(link = "log"))
glm_mod5_pred <- predict(glm_mod5, newdata = df, type = "response", se.fit = TRUE)

glm_mod6 <- glm(nuovi_positivi ~ (ricoverati_con_sintomi + terapia_intensiva) * color + data + ricoverati_con_sintomi_prev + nuovi_tamponi_pcr, data = df, family=poisson(link = "log"))
glm_mod6_pred <- predict(glm_mod6, newdata = df, type = "response", se.fit = TRUE)

summary(glm_mod5)
summary(glm_mod6)

ggplot() +
  geom_point(data = df, aes(x = data, y = nuovi_positivi, color="nuovi positivi", group = 1),col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))])+ 
  geom_line(data = df, aes(x = data, y = predict(glm_mod5,type="response"), color = "model 5", group = 1)) +
  geom_line(data = df, aes(x = data, y = predict(glm_mod6,type="response"), color = "model 6", group = 1)) +
  geom_ribbon(data=df, aes(x=data, ymin=glm_mod6_pred$fit - glm_mod6_pred$se.fit, ymax=glm_mod6_pred$fit + glm_mod6_pred$se.fit),alpha=0.5)+
  geom_ribbon(data=df, aes(x=data, ymin=glm_mod5_pred$fit - glm_mod5_pred$se.fit, ymax=glm_mod5_pred$fit + glm_mod5_pred$se.fit),alpha=0.5)+
  
  labs(x = "Time", y = "New positives") +
  scale_x_date(date_breaks = '5 days',
               date_labels = '%Y-%m-%d') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

AIC(glm_mod5, glm_mod6)


library(modelr)

df$resid <- df$nuovi_positivi - predict(glm_mod6,type="response")

df %>% 
  ggplot(aes(data, resid)) + 
  geom_ref_line(h = 0) + 
  geom_line(col=1) + 
  geom_smooth(se = FALSE, span = 0.20, col=3)

```

```{r glm_mod6 predictions, echo=FALSE, message=FALSE, warning=FALSE}

glm_mod6_pred <- predict(glm_mod6, newdata = df_extended, type = "response", se.fit = TRUE)

ggplot() +
  geom_point(data = df_extended, aes(x = data, y = nuovi_positivi, color="nuovi positivi", group = 1),col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df_extended$color))])+ 
  geom_line(data = df_extended, aes(x = data, y = predict(glm_mod6, newdata=df_extended, type="response"), color = "model 4", group = 1)) +  geom_ribbon(data=df_extended, aes(x=data, ymin=glm_mod6_pred$fit - glm_mod6_pred$se.fit, ymax=glm_mod6_pred$fit + glm_mod6_pred$se.fit),alpha=0.5)+
  
  
  labs(x = "Time", y = "New positives") +
  scale_x_date(date_breaks = '5 days',
               date_labels = '%Y-%m-%d') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

```{r glm_mod7, echo=FALSE, message=FALSE, warning=FALSE}

glm_mod7 <- glm(nuovi_positivi ~  (ricoverati_con_sintomi + terapia_intensiva + ricoverati_con_sintomi_prev + nuovi_tamponi_pcr) * color + data , data = df, family=poisson(link = "log"))

glm_mod7_pred <- predict(glm_mod7, newdata = df, type = "response", se.fit = TRUE)

summary(glm_mod7)

ggplot() +
  geom_point(data = df, aes(x = data, y = nuovi_positivi, color="nuovi positivi", group = 1),col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))])+ 
  geom_line(data = df, aes(x = data, y = predict(glm_mod7,type="response"), color = "model 4", group = 1)) +
  geom_ribbon(data=df, aes(x=data, ymin=glm_mod7_pred$fit - glm_mod7_pred$se.fit, ymax=glm_mod7_pred$fit + glm_mod7_pred$se.fit),alpha=0.5)+
  
  
  labs(x = "Time", y = "New positives") +
  scale_x_date(date_breaks = '5 days',
               date_labels = '%Y-%m-%d') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

AIC(glm_mod7,glm_mod6,glm_mod5,glm_mod4,glm_mod3,glm_mod2,glm_mod1)
```


```{r glm_mod7 predictions, echo=FALSE, message=FALSE, warning=FALSE}

glm_mod7_pred <- predict(glm_mod7, newdata = df_extended, type = "response", se.fit = TRUE)

mean(glm_mod7_pred$se.fit[])/mean(df_extended$nuovi_positivi)

ggplot() +
  geom_point(data = df_extended, aes(x = data, y = nuovi_positivi, color="nuovi positivi", group = 1),col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df_extended$color))])+ 
  geom_line(data = df_extended, aes(x = data, y = predict(glm_mod7, newdata=df_extended, type="response"), color = "model 4", group = 1)) +
  geom_ribbon(data=df_extended, aes(x=data, ymin=glm_mod7_pred$fit - glm_mod7_pred$se.fit, ymax=glm_mod7_pred$fit + glm_mod7_pred$se.fit),alpha=0.5)+
  
  
  labs(x = "Time", y = "New positives") +
  scale_x_date(date_breaks = '5 days',
               date_labels = '%Y-%m-%d') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

## GAM
Since the best glm model is the last one (from the AIC ), I will try to build a GAM with same covariates.

```{r gam_mod1, echo=FALSE}
library(mgcv)
gam_mod1 <- gam(nuovi_positivi ~  (ricoverati_con_sintomi + terapia_intensiva + ricoverati_con_sintomi_prev + nuovi_tamponi_pcr) * color + data , data = df )
gam_mod1_pred <- predict(gam_mod1, type="response", se.fit = TRUE)
summary(gam_mod1)

ggplot() +
  geom_point(data = df, aes(x = data, y = nuovi_positivi, color="nuovi positivi", group = 1),col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))])+ 
  geom_line(data = df, aes(x = data, y = predict(gam_mod1,type="response"), color = "model 4", group = 1)) +
  geom_ribbon(data=df, aes(x=data, ymin=gam_mod1_pred$fit - gam_mod1_pred$se.fit, ymax=gam_mod1_pred$fit + gam_mod1_pred$se.fit),alpha=0.2)+
  
  
  labs(x = "Time", y = "New positives") +
  scale_x_date(date_breaks = '5 days',
               date_labels = '%Y-%m-%d') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r gam predictions, echo=FALSE, message=FALSE, warning=FALSE}
gam_mod_1_pred <- predict(gam_mod1, newdata=df_extended, type="response", se.fit = TRUE)
mean(gam_mod_1_pred$se.fit)
ggplot() +
  geom_point(data = df_extended, aes(x = data, y = nuovi_positivi, color="nuovi positivi", group = 1),col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df_extended$color))])+ 
  geom_line(data = df_extended, aes(x = data, y = predict(gam_mod1, newdata=df_extended, type="response"), color = "model 4", group = 1)) +
  geom_ribbon(data = df_extended, aes(x= data, ymin=-gam_mod_1_pred$se.fit + gam_mod_1_pred$fit, ymax=gam_mod_1_pred$se.fit + gam_mod_1_pred$fit), alpha=0.5)+
  
  
  labs(x = "Time", y = "New positives") +
  scale_x_date(date_breaks = '5 days',
               date_labels = '%Y-%m-%d') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

## LOESS
```{r LOESS}
library(mgcv)
loess_mod1 <- loess(nuovi_positivi ~ terapia_intensiva + ricoverati_con_sintomi_prev + nuovi_tamponi_pcr  , data = df )
loess_mod1_pred <- predict(loess_mod1, type="response", se.fit = TRUE)
summary(loess_mod1)

ggplot() +
  geom_point(data = df, aes(x = data, y = nuovi_positivi, color="nuovi positivi", group = 1),col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))])+ 
  geom_line(data = df, aes(x = data, y = predict(loess_mod1,type="response"), color = "model 4", group = 1)) +
  
  
  
  labs(x = "Time", y = "New positives") +
  scale_x_date(date_breaks = '5 days',
               date_labels = '%Y-%m-%d') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r loess predictions, echo=FALSE, message=FALSE, warning=FALSE}

ggplot() +
  geom_point(data = df_extended, aes(x = data, y = nuovi_positivi, color="nuovi positivi", group = 1),col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df_extended$color))])+ 
  geom_line(data = df_extended, aes(x = data, y = predict(loess_mod1,  newdata=df_extended, type="response"), color = "model 4", group = 1)) +
  
  
  
  labs(x = "Time", y = "New positives") +
  scale_x_date(date_breaks = '5 days',
               date_labels = '%Y-%m-%d') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

