---
title: "Covid-19 - SMSD"
output:
  html_document: default
---
  
#### Goal
  
Build up a useful model in order to predict response variable Y starting from data provided.

-   response variable: number of new positives
-   data: Covid-19 spreading outbreak from the official website of Protezione Civile

#### Dataset

Load dataset.

Read data we are interested in, then those that are relative to the following couple of parameters:
  
-   regione: Sicilia

-   data: 1st October 2020 to 1st February 2021

```{r dataset, message=FALSE, echo = FALSE}
df_extended <- data.frame(read.csv("sicily_secondwave_covid.csv"))
df_extended$data <- as.Date(df_extended$data,  "%Y-%m-%d")

df <- df_extended[which(df_extended$data >= "2020-10-01" & df_extended$data <= "2021-01-31"),]
df$data <- as.Date(df$data,  "%Y-%m-%d")

col_zone <- c("#fc6b03","#cfcaca","#f2d729","#b3190b")

```

#### Quality of the data

Check for missing values in the data frame

```{r na, echo=FALSE}
ncols <- dim(df)[2]
na_count <- sapply(df[,-c((ncols-15):ncols)], function(y) sum(length(which(is.na(y)))))

na_count <- data.frame( gsub("_", " ", names(na_count)),na_count)
colnames(na_count) <- c("variable", "missing values")

library(gridExtra)
library(grid)
color_df <- data.frame(color = c("#bed0e8", "#809cc2"), stringsAsFactors = FALSE)
my_table_theme <- ttheme_default(core=list(bg_params = list(fill = color_df$color[1:2], col=NA),padding=unit.c(unit(4, "mm"), unit(10, "mm"))), colhead =list(bg_params=list(fill ="#5680ba")))


grid.newpage()
grid.draw(tableGrob(na_count, rows=NULL))

# png("images/missing_values_table.png", width=4.5,height=6.5,bg = "white",units="in",res=1200)
# grid.table(na_count, rows=NULL)
# dev.off()


```

Missing value are about the admission to the ICU and for this reason that value has been removed and it has been considered the number of ICU per day.

#### Clean dataset

Get rid of redundant and not useful variables, substitute or replace missing values and finally check for any kinf of relationship between variables.

```{r plot response var against time, echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(patchwork)

ggplot() +
  geom_line(data = df, aes(x = data, y = nuovi_positivi, group = 1)) +
  geom_point(data = df, aes(x = data, y = nuovi_positivi, group = 1),col=col_zone[unclass(as.factor(df$color))], size=2)+
  labs(x = "Time", y = "New positives") +
  scale_x_date(date_breaks = '5 days',
               date_labels = '%Y-%m-%d') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
ggsave(paste0( "images/response_variable.png"), width = 20, height = 8, dpi = 150)


```

```{r prior analysis, echo=FALSE, message=FALSE, warning=FALSE}
library("ggcorrplot")
library("corrplot")

corr_df <- df
corr_df$data <- as.numeric(corr_df$data)
cols <- c("data", "ricoverati_con_sintomi","terapia_intensiva","totale_ospedalizzati", "isolamento_domiciliare", "totale_positivi", "dimessi_guariti", "deceduti", "tamponi", "totale_casi", "casi_testati","nuovi_positivi")

M=cor(corr_df[,cols])
colnames(M) <- c("A", "B", "C", "D", "E", "F", "G", "H", "I", "L", "K", "I")
rownames(M) <- paste0(colnames(M), ". ", gsub("_", " ", cols))

col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))

corrplot(M,col.lim=c(-1,1),is.corr=FALSE, method="color", col=col(200), type="lower", addCoef.col = "black", tl.col = "black",tl.srt = 360, tl.offset = 1, tl.cex=0.7)

# png("images/variables_corrplot.png", width=10,height=8,units="in",res=1500)
# corrplot(M,col.lim=c(0,1),is.corr=FALSE, method="color",col=col(200),
#      type="lower", addCoef.col = "black", tl.col = "black",tl.srt = 360, tl.offset = 1, tl.cex=1.1)
# dev.off()


```
<br>
  
  We can see there are many variables which are highly related, though this relationship can be due to the fact that they are not independent. While for others variables we observe a poor correlation while we would have expected something different, for example for the number of swabs carried out.
For this reason new variables were introduced for those values which account for the total amount of cases since the beginning of the pandemic.

-   nuovi_tamponi: swabs carried out daily (tamponi current day - tamponi predious day) [by the 15/01 were counted also the antigenic tests]

-   nuovi_decessi: daily deaths (deceduti current day - deceduti previous day)

#### Perform some explanatory analysis of the data

Then a new correlation plot is made to show which are the most highly correlated variables.

```{r correlation on chosen covariates, echo=FALSE, message=FALSE, warning=FALSE}
library("ggcorrplot")
library("corrplot")
library("dvmisc")
df$nuovi_tamponi <- df$nuovi_tamponi_pcr
cols <- c("ricoverati_con_sintomi", "isolamento_domiciliare","terapia_intensiva", "nuovi_tamponi",  "nuovi_positivi")
M=cor(df[,cols])
colnames(M) <- c("A", "B", "C", "D", "E")
rownames(M) <- paste0(colnames(M), ". ", gsub("_", " ", cols))

col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))

corrplot(M, method="color", col=col(200),  col.lim=c(0,1), is.corr=FALSE, type="lower", addCoef.col = "black", tl.col = "black", tl.srt = 360, tl.offset = 1, tl.cex=0.7)

# png("images/variables_corrplot2.png", width=10,height=8,units="in",res=1500)
# corrplot(M,col.lim=c(0,1),is.corr=FALSE, method="color",col=col(200),
#      type="lower", addCoef.col = "black", tl.col = "black",tl.srt = 360, tl.offset = 1, tl.cex=1.1)
# dev.off()

```

<br>
  Further investigate the correlations found between each pair of variables

```{r scatterplot covariates, echo=FALSE}
#png("images/scatterplot_covariates.png", width=13,height=6.5,bg = "white",units="in",res=1500)

par(mfrow=c(2,3))
plot(ricoverati_con_sintomi ~ terapia_intensiva, data = df, pch =16, xlab = "terapia intensiva", ylab = "ricoverati con sintomi", col="#0b084d")
plot(ricoverati_con_sintomi ~ nuovi_tamponi_pcr, data = df, pch =16, xlab = "nuovi tamponi", ylab = "ricoverati con sintomi", col="#0b084d")
plot(ricoverati_con_sintomi ~ isolamento_domiciliare, data = df, pch =16, xlab = "isolamento domiciliare", ylab = "ricoverati con sintomi", col="#0b084d")

plot(terapia_intensiva ~ nuovi_tamponi_pcr, data = df, pch =16, xlab = "nuovi tamponi", ylab = "terapia intensiva", col = "#0b084d")

plot(isolamento_domiciliare ~ nuovi_tamponi_pcr, data = df, pch =16, xlab = "nuovi tamponi", ylab = "isolamento domiciliare", col="#0b084d")
plot(isolamento_domiciliare ~ terapia_intensiva, data = df, pch =16, xlab = "terapia intensiva", ylab = "isolamento domiciliare", col="#0b084d")

#dev.off()
```

The following plot shows the trends of previous variable through all the time period considered, it can be seen how it is almost the same for each of them, exept for the number of swabs.

```{r most important variables trend, echo=FALSE, warning = FALSE}
p1 <- ggplot() +
  geom_line(data = df, aes(x = data, y = log(nuovi_positivi), color="nuovi positivi", group = 1),ylab="") +
  geom_point(data = df, aes(x = data, y = log(nuovi_positivi), group = 1))+ 
  geom_line(data = df, aes(x = data, y = log(nuovi_tamponi_pcr), color="nuovi tamponi", group = 1)) +
  geom_point(data = df, aes(x = data, y = log(nuovi_tamponi_pcr),color="nuovi tamponi", group = 1)) +
  geom_line(data = df, aes(x = data, y = log(ricoverati_con_sintomi), color="ricoverati con sintomi", group = 1)) +
  geom_point(data = df, aes(x = data, y = log(ricoverati_con_sintomi),color="ricoverati con sintomi", group = 1)) +
  geom_line(data = df, aes(x = data, y = log(terapia_intensiva), color="terapia intensiva", group = 1)) +
  geom_point(data = df, aes(x = data, y = log(terapia_intensiva),color="terapia intensiva", group = 1)) +
  geom_line(data = df, aes(x = data, y = log(isolamento_domiciliare), color="isolamento domiciliare", group = 1)) +
  geom_point(data = df, aes(x = data, y = log(isolamento_domiciliare),color="isolamento domiciliare", group = 1)) +
  scale_x_date(date_breaks = '5 days', date_labels = '%Y-%m-%d') +
  labs(y = "log") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_colour_manual(values = c("nuovi positivi" = "black", "nuovi tamponi" ="#2f6d80", "ricoverati con sintomi" = "#490e96", "terapia intensiva" = "#1a8f2f", "isolamento domiciliare" = "#851128")) 

p1
#ggsave(paste0( "images/variables_trend.png"), plot=p1, height=6, width=8, units="in", dpi = 150)
```



Previous chosen covariates were part of the given dataset, now we investigate possible relationships with information which is not part of our dataset.
The first covariate which has been considered and added is related to the restrictions to which the region was subjected in the period considered. These restrictions are of 4 different levels, which are the 4 possible regional colors: white, yellow, orange, red.

-   See how response variable changes wrt to region colors:
  
```{r boxplot_scatterplot, echo =FALSE, warning = FALSE}
library(ggplot2)
library(RColorBrewer)

col <- c("#FF9636","#F9F6F0","#FFCD58","#A82810")

ggplot(df,aes(x=factor(color),y=nuovi_positivi))+
  geom_boxplot(outlier.color="black", fill=col) +
  labs(x = " ", y = "nuovi positivi")+
  geom_jitter(color="black", width = 0) 

#ggsave(paste0("images/boxplot_color.png"), width = 8, height = 5, dpi = 200)

p2 <- ggplot(df,aes(x=factor(color),y=ricoverati_con_sintomi)) +
  geom_boxplot(outlier.color="black", fill=col) +
  labs(x = " ", y = "ricoverati con sintomi") +
  geom_jitter(color="black", width = 0) 

p3 <- ggplot(df,aes(x=factor(color),y=nuovi_tamponi_pcr)) +
  geom_boxplot(outlier.color="black", fill=col) +
  labs(x = " ", y = "nuovi tamponi") +
  geom_jitter(color="black", width = 0) 

p2
#ggsave(paste0("images/boxplot_color1.png"), width = 8, height = 5, dpi = 200)

p3
#ggsave(paste0("images/boxplot_color2.png"), width = 8, height = 5, dpi = 200)

ggplot(df, aes(x=weekdays(df$data), y=df$nuovi_tamponi_pcr)) + 
  geom_boxplot(outlier.color="black", fill=c("#E7F2F8","#74BDCB","#B5E5CF","#EFE7BC","#E98973","#B99095","#88B2CC")) +
  labs(x = "giorno della settimana", y = "numero tamponi") +
  geom_jitter(color="black", width = 0) +
   theme(axis.title.x = element_text(size = rel(1.2)), axis.title.y = element_text(size = rel(1.2)),
        title = element_text(size = rel(1.2)))

#ggsave(paste0("images/boxplot_wekdays_tamponi.png"), width = 16, height = 9, dpi = 300)

ggplot(df, aes(x=weekdays(df$data), y=df$nuovi_positivi)) + 
  geom_boxplot(outlier.color="black", fill=c("#E7F2F8","#74BDCB","#B5E5CF","#EFE7BC","#E98973","#B99095","#88B2CC")) +
  labs(x = "giorno della settimana", y = "nuovi positivi") +
  geom_jitter(color="black", width = 0) +
   theme(axis.title.x = element_text(size = rel(1.2)), axis.title.y = element_text(size = rel(1.2)),
        title = element_text(size = rel(1.2)))

#ggsave(paste0("images/boxplot_wekdays_positivi.png"), width = 16, height = 9, dpi = 300)

```


```{r variables, echo=FALSE, message=FALSE, warning = FALSE, comment = FALSE}
#png("images/corr_scatterplot_response_variable.png", width=10,height=10,units="in",res=1200)
par(mfrow=c(2,2))
plot(nuovi_positivi ~ ricoverati_con_sintomi, data = df, pch =16, xlab = "ricoverati con sintomi", ylab = "nuovi positivi",col=col_zone[unclass(as.factor(df$color))])
plot(nuovi_positivi ~ terapia_intensiva, data = df, pch =16, xlab = "terapia intensiva", ylab = "nuovi positivi", col=col_zone[unclass(as.factor(df$color))])
plot(nuovi_positivi ~ isolamento_domiciliare, data = df, pch =16, xlab = "isolamento domiciliare", ylab = "nuovi positivi",col=col_zone[unclass(as.factor(df$color))])
plot(nuovi_positivi ~ nuovi_tamponi_pcr, data = df, pch =16, xlab = "nuovi tamponi", ylab = "nuovi positivi",col=col_zone[unclass(as.factor(df$color))])

#dev.off()

```


## Building models

We start building our model looking at the relationship between our response variable and its highest correlated predictor (which is given by people hospitalized with symptoms, maybe because the highest the number of people hospitalized, the higher the contagiousness and as a consequence the higher the number of new positives).

### Linear models

```{r linear models, echo=FALSE, message=FALSE, warning=FALSE}
# Try some linear models to see which are the variables which significantly explain the variability of y

formula <- paste0("nuovi_positivi ~ ", c("ricoverati_con_sintomi", 
                                         "ricoverati_con_sintomi + color", 
                                         "ricoverati_con_sintomi * color", 
                                         "ricoverati_con_sintomi + color + nuovi_tamponi_pcr", 
                                         "(ricoverati_con_sintomi + nuovi_tamponi_pcr) * color"))

lm_mod1 <- lm(eval(formula[1]), data = df)
lm_mod2 <- lm(eval(formula[2]), data = df)
lm_mod3 <- lm(eval(formula[3]), data = df)
lm_mod4 <- lm(eval(formula[4]), data = df)
lm_mod5 <- lm(eval(formula[5]), data = df)

library("RColorBrewer")

# summary(lm_mod1)
# summary(lm_mod2)
# summary(lm_mod3)
# summary(lm_mod4)
# summary(lm_mod5)

ggplot() +
  geom_point(data = df, aes(x = data, y = nuovi_positivi, color="nuovi positivi", group = 1),col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))])+ 
  geom_line(data = df, aes(x = data, y = predict(lm_mod3), color = "model 3", group = 1), size=0.75) +
  geom_line(data = df, aes(x = data, y = predict(lm_mod4), color = "model 4", group = 1), size=0.75) +
  geom_line(data=df, aes(x = data, y = predict(lm_mod5), color = "model 5"), size=0.75)+
  
  labs(x = "Time", y = "New positives") +
  scale_x_date(date_breaks = '5 days',
               date_labels = '%Y-%m-%d') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_color_brewer(palette = "Dark2")
ggsave(paste0("images/lm_mod.png"), width = 12, height = 5, dpi = 200)


```


```{r lm table,  echo=FALSE, message=FALSE, warning=FALSE}
lm_aic <- AIC(lm_mod1,lm_mod2,lm_mod3,lm_mod4, lm_mod5)
#lm_mselm<-rbind(get_mse(lm_mod1),get_mse(lm_mod2),get_mse(lm_mod3),get_mse(lm_mod4),get_mse(lm_mod5))


lm_table <- data.frame(formula,lm_aic[2])
colnames(lm_table) <- c("Model", "AIC")

n<- dim(lm_table)[1]

tt3 <- ttheme_minimal(
  core=list(bg_params = list(fill = blues9[1:2], col=NA), padding=unit.c(unit(15, "mm"), unit(10, "mm")),fg_params=list(hjust=0, x=c(rep(0.02,n), rep(0.2,n)))),
  colhead=list(bg_params=list(fill="#cee4f5"),fg_params=list(col="#0c1b4a",fontsize=14, fontface="bold",hjust=0, x=c(0.02,0.3)),padding=unit.c(unit(4, "mm"), unit(10, "mm"))),
  rowhead=list(fg_params=list(col="orange", fontface=3L, hjust=0, x=0)))

# png("images/lm_table.png", width=10,height=8,units="in",res=1500)
# grid.newpage()
# grid.table(lm_table, rows = NULL, theme = tt3)
# dev.off()
```

<br>

### Generalized linear models

#### Poisson distirbution

```{r glm, echo=FALSE, message=FALSE, warning=FALSE}

glm_mod1 <- glm(eval(formula[1]), data = df, family=poisson(link = "log"))
glm_mod2 <- glm(eval(formula[2]), data = df, family=poisson(link = "log"))
glm_mod3 <- glm(eval(formula[3]), data = df, family=poisson(link = "log"))
glm_mod4 <- glm(eval(formula[4]), data = df, family=poisson(link = "log"))
glm_mod5 <- glm(eval(formula[5]), data = df, family=poisson(link = "log"))

# plot(glm_mod1)
# plot(glm_mod2)
# plot(glm_mod3)
# plot(glm_mod4)
# plot(glm_mod5)

ggplot() +
  geom_point(data = df, aes(x = data, y = nuovi_positivi, color="nuovi positivi", group = 1),col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))])+ 
  geom_line(data = df, aes(x = data, y = predict(glm_mod3,type="response"), color = "model 3", group = 1), size=0.75) +
  geom_line(data = df, aes(x = data, y = predict(glm_mod4,type="response"), color = "model 4", group = 1),size=0.75) +
  geom_line(data= df, aes(x= data, y = predict(glm_mod5, type="response"), color="model 5"), size=0.75)+
  labs(x = "Time", y = "New positives") +
  scale_x_date(date_breaks = '5 days',
               date_labels = '%Y-%m-%d') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_color_brewer(palette = "Dark2")

#ggsave(paste0("images/glm_mod.png"), width = 12, height = 5, dpi = 200)

AIC(glm_mod1, glm_mod2, glm_mod3, glm_mod4, glm_mod5)

```

Check residual plot with respect to region color:

```{r glm residuals, echo =FALSE, message=FALSE, warning=FALSE}
## Model checking - Residual plots
library(modelr)
library(ggplot2)

df$resid <- df$nuovi_positivi -  predict(glm_mod4,type="response")

ggplot(df, aes(data, resid, colour = color)) + 
  geom_ref_line(h = 0) + 
  geom_line()

df %>% 
  ggplot(aes(data, resid)) + 
  geom_ref_line(h = 0) + 
  geom_line(col=1) + 
  geom_smooth(se = FALSE, span = 0.20, col=3)

pred <- predict(glm_mod5, type="response")
res <- residuals(glm_mod5, type="deviance")

ggplot() +
  geom_point(aes(x = log(pred), y = res), col="#2783c4") +
  geom_smooth(aes(x = log(pred), y = res), col ="#45a377") +
  geom_line(aes(x= log(pred), y = 0), lty=2) +
  labs(x = "Fitted values", y = "Residuals", title = "Linearity", subtitle ="Rference line should be flat and horizontal") +
  theme_light() +
  theme(axis.title.x = element_text(size = rel(1.1)),axis.title.y = element_text(size = rel(1.1)),
        title = element_text(size = rel(1.3)))

#ggsave(paste0("images/residuals_glm_mod.png"), width = 10, height = 6, dpi = 200)


```

```{r glm table,  echo=FALSE, message=FALSE, warning=FALSE}
glm_aic <- AIC(glm_mod1,glm_mod2,glm_mod3,glm_mod4, glm_mod5)
glm_mselm<-rbind(get_mse(glm_mod1),get_mse(glm_mod2),get_mse(glm_mod3),get_mse(glm_mod4),get_mse(glm_mod5))

glm_table <- data.frame(formula,glm_aic[2], glm_mselm)
colnames(glm_table) <- c("Model", "AIC", "MSE")


#library(gtable)
#png("images/glm_table.png", width=10,height=8,units="in",res=1500)
#grid.newpage()
#grid.table(glm_table, rows = NULL, theme = tt3)
#dev.off()
```

#### GLM with negative binomial

```{r nglm, echo=FALSE, message = FALSE, warning=FALSE, comment=FALSE}

#GLM with negative binomial
require("MASS")
nrw<-length(df[,1]) 

listlm<-c("nglmod1","nglmod2","nglmod3","nglmod4","nglmod5")

param <-  c(" ricoverati_con_sintomi", 
             " ricoverati_con_sintomi + color", 
             " ricoverati_con_sintomi * color", 
             " ricoverati_con_sintomi + color + nuovi_tamponi_pcr", 
             " (ricoverati_con_sintomi + nuovi_tamponi_pcr) * color")
num_var<-5
l<-colnames(df)
counter<-1
str<-"nuovi_positivi"

for (i in 1:num_var){
   if(counter>num_var) break
  name<-paste("nglmod",counter, sep="")
  f<-as.formula(paste(str, param[counter] ,sep = " ~ "))
  tmp<-glm.nb(formula=f, data = df)
  assign(name,tmp) 
  counter = counter +1
}

Gplot<-ggplot() +
     geom_point(data = df, aes(x = data, y = nuovi_positivi, color="nuovi positivi", group = 1),col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))])+ scale_x_date(date_breaks = '5 days',
               date_labels = '%Y-%m-%d') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  labs(x = "Time", y = "New positives") 
    counter<-1
    tmpdf<-NULL
    boh<-rep(0,nrw)
    y<-matrix(nrow=nrw, ncol=num_var)
    tmp<-nrw*num_var
    m<-rep(0,tmp)
    for(i in listlm){
      name<-paste("model",counter,sep="")
      tmp1<-paste("tmpdf",counter,sep="")
        tmp<-as.name(i)
      y[,counter]<-predict(eval(tmp),type="response")

    l<-(nrw*(counter-1))
    if(counter==1) l=1
    r<-(nrw*counter)
    m[l:r]<-name
    counter<-counter+1
}
    
y<-as.vector(y)
x<-rep(df$data,num_var)
dfn<-data.frame("x"=x,"y"=y,"model"=m)
Gplot<-Gplot+geom_line( data=dfn,aes(x = x, y = y, color=model, group =model)) 
AIC(nglmod1,nglmod2,nglmod3,nglmod4)



AICn <- rbind(extractAIC(nglmod1)[2],extractAIC(nglmod2)[2],
             extractAIC(nglmod3)[2])
#BIC <- rbind(extractAIC(lm1, k=log(n))[2],extractAIC(lm2,k=log(n))[2],
 #            extractAIC(lm3,k=log(n))[2],extractAIC(lm4,k=log(n))[2])

# plot models
ggplot() +
  geom_point(data = df, aes(x = data, y = nuovi_positivi, color="nuovi positivi", group = 1),col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))])+ 
  geom_line(data = df, aes(x = data, y = predict(nglmod3,type="response"), color = "model 3", group = 1), size=0.75) +
  geom_line(data = df, aes(x = data, y = predict(nglmod4,type="response"), color = "model 4", group = 1),size=0.75) +
  geom_line(data= df, aes(x= data, y = predict(nglmod5, type="response"), color="model 5"), size=0.75)+
  
  
  labs(x = "Time", y = "New positives") +
  scale_x_date(date_breaks = '5 days',
               date_labels = '%Y-%m-%d') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_color_brewer(palette = "Dark2")

#ggsave(paste0("images/nglm_mod.png"), width = 12, height = 5, dpi = 200)


# plot residuals

pred <- predict(nglmod5, type="response")
res <- residuals(nglmod5, type="deviance")

ggplot() +
  geom_point(aes(x = log(pred), y = res), col="#2783c4") +
  geom_smooth(aes(x = log(pred), y = res), col ="#45a377") +
  geom_line(aes(x = log(pred), y = 0), lty=2) +
  labs(x = "Fitted values", y = "Residuals", title = "Linearity", subtitle ="Reference line should be flat and horizontal") +
  theme_light() +
  theme(axis.title.x = element_text(size = rel(1.1)),axis.title.y = element_text(size = rel(1.1)),
        title = element_text(size = rel(1.3)))

ggsave(paste0("images/residuals_nglm_mod.png"), width = 10, height = 6, dpi = 200)

```


```{r ngl table, echo = FALSE, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
ngl_aic <- AIC(nglmod1,nglmod2,nglmod3,nglmod4, nglmod5)

ngl_mselm<-rbind(mean(resid(nglmod1)^2),mean(resid(nglmod2)^2),mean(resid(nglmod3)^2),mean(resid(nglmod4)^2),mean(resid(nglmod5)^2))

ngl_table <- data.frame(formula,ngl_aic[2], ngl_mselm)
colnames(ngl_table) <- c("Model", "AIC", "MSE")


library(gtable)
#png("images/nglm_table.png", width=10,height=8,units="in",res=1500)
grid.newpage()
grid.table(ngl_table, rows = NULL, theme = tt3)
#dev.off()
```

# Performance tab


```{r performance table, echo = FALSE,  message=FALSE, warning=FALSE, include=FALSE}
library(performance)
library(gtable)
library(grid)
library(gridExtra)

formula <- paste0("nuovi_positivi ~ ", c("(ricoverati_con_sintomi + nuovi_tamponi_pcr) * color" , "ricoverati_con_sintomi + color + nuovi_tamponi_pcr","ricoverati_con_sintomi * color","ricoverati_con_sintomi + color" ,"ricoverati_con_sintomi" ))


all_compare <- compare_performance(lm_mod1,lm_mod2,lm_mod3,lm_mod4,lm_mod5,glm_mod1,glm_mod2,glm_mod3,glm_mod4,glm_mod5,nglmod1,nglmod2,nglmod3,nglmod4,nglmod5, rank=TRUE)
all_compare
all_compare$Name <- c(formula,formula,formula)

lm_compare <- compare_performance(lm_mod1,lm_mod2,lm_mod3,lm_mod4,lm_mod5, rank=TRUE)
lm_aic <- AIC(lm_mod5,lm_mod4,lm_mod3,lm_mod2, lm_mod1)
lm_compare$AIC <- lm_aic$AIC
lm_compare
lm_compare$Name <- formula



glm_compare <- compare_performance(glm_mod1,glm_mod2,glm_mod3,glm_mod4,glm_mod5, rank=TRUE)
glm_compare <- glm_compare[,-c(6,7)]
glm_aic <- AIC(glm_mod5,glm_mod4,glm_mod3,glm_mod2, glm_mod1)
glm_compare$R2_Nagelkerke <- glm_aic$AIC
colnames(glm_compare)[3] <- "AIC" 
glm_compare$Name <- formula


ngb_compare <- compare_performance(nglmod1, nglmod2,nglmod3,nglmod4,nglmod5, rank=TRUE)
ngb_compare <- ngb_compare[,-c(6,7)]
ngb_aic <- AIC(nglmod5,nglmod4,nglmod3,nglmod2, nglmod1)
ngb_compare$R2_Nagelkerke <- ngb_aic$AIC
colnames(ngb_compare)[3] <- "AIC" 
ngb_compare$Name <- formula

ngb_compare

# png("images/all_compare_table.png", width=28,height=9,units="in",res=1000)
# grid.newpage()
# grid.table(all_compare, rows = NULL, theme = tt3)
# dev.off()

# png("images/lm_compare_table.png", width = 20, height = 8, units = "in", res=1000)
# grid.newpage()
# grid.table(lm_compare,rows = NULL, theme = tt3)
# dev.off()
# 
# png("images/glm_compare_table.png", width=20,height=8,units="in",res=1000)
# grid.newpage()
# grid.table(glm_compare, rows = NULL, theme = tt3)
# dev.off()

#png("images/ngb_compare_table.png", width=20,height=8,units="in",res=1000)
#grid.newpage()
#grid.table(ngb_compare, rows = NULL, theme = tt3)
#dev.off()

```





## LOESS

```{r LOESS, echo = FALSE, message=FALSE, warning=FALSE}
library(mgcv)
loess_mod1 <- loess(nuovi_positivi ~ terapia_intensiva + ricoverati_con_sintomi_prev + nuovi_tamponi_pcr  , data = df )
loess_mod1_pred <- predict(loess_mod1, type="response", se.fit = TRUE)
summary(loess_mod1)

d<-df
d$data <- as.numeric(d$data)

loess_mod1<-loess(nuovi_positivi~ data + ricoverati_con_sintomi + nuovi_tamponi_pcr, data = d)

ggplot() +
  geom_point(data = df, aes(x = data, y = nuovi_positivi, color="nuovi positivi", group = 1),col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))])+ 
  geom_line(data = df, aes(x = data, y = predict(loess_mod1,type="response"), color = "loess model", group = 1)) +
  labs(x = "Time", y = "New positives") +
  scale_x_date(date_breaks = '5 days',
               date_labels = '%Y-%m-%d') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```



# Plot response on covariates

### With respect to best glm model:
```{r plot, echo = FALSE, message=FALSE, warning=FALSE}
ggplot(data=df)+
  geom_point(aes(x=ricoverati_con_sintomi, y=nuovi_positivi), col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))])+
  geom_line(aes(x=ricoverati_con_sintomi,y=predict(glm_mod5, type="response"),color = "Fit", group = 1), size=0.75)+
  scale_color_brewer(palette = "Dark2")
#ggsave(paste0("images/glm_mod_ric_vs_Y.png"), width = 12, height = 5, dpi = 200)

ggplot(data=df)+
  geom_point(aes(x=nuovi_tamponi_pcr, y=nuovi_positivi), col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))])+
  geom_line(aes(x=nuovi_tamponi_pcr,y=predict(glm_mod5, type="response"),color = "Fit", group = 1), size=0.75)+
  scale_color_brewer(palette = "Dark2")
#ggsave(paste0("images/glm_mod_pcr_vs_Y.png"), width = 12, height = 5, dpi = 200)

ggplot(data=df)+
  geom_point(aes(x=ricoverati_con_sintomi, y=nuovi_positivi), col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))])+
  geom_line(aes(x=ricoverati_con_sintomi,y=predict(lm_mod5, type="response"),color = "Fit", group = 1), size=0.75)+
  scale_color_brewer(palette = "Dark2")

#ggsave(paste0("images/lm_mod_ric_vs_Y.png"), width = 12, height = 5, dpi = 200)

ggplot(data=df)+
  geom_point(aes(x=nuovi_tamponi_pcr, y=nuovi_positivi), col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))])+
  geom_line(aes(x=nuovi_tamponi_pcr,y=predict(lm_mod5, type="response"),color = "Fit", group = 1), size=0.75)+
  scale_color_brewer(palette = "Dark2")

#ggsave(paste0("images/lm_mod_pcr_vs_Y.png"), width = 12, height = 5, dpi = 200)


```

## GAM
Since the best glm model is the last one (from the AIC ), I will try to build a GAM using the same covariates.

```{r gam_mod1, echo=FALSE, message=FALSE, warning=FALSE}
library(mgcv)
gam_mod1 <- gam(nuovi_positivi ~  s(ricoverati_con_sintomi*color) + s(nuovi_tamponi_pcr*color), data = df,family=Gamma(link=log) )

gam_mod1_pred <- predict(gam_mod1, type="response", se.fit = TRUE)
summary(gam_mod1)

ggplot() +
  geom_point(data = df, aes(x = data, y = nuovi_positivi, color="nuovi positivi", group = 1),col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))])+ 
  geom_line(data = df, aes(x = data, y = predict(gam_mod1,type="response"), color = "model 4", group = 1)) +
  geom_ribbon(data=df, aes(x=data, ymin=gam_mod1_pred$fit - gam_mod1_pred$se.fit, ymax=gam_mod1_pred$fit + gam_mod1_pred$se.fit),alpha=0.2)+
  labs(x = "Time", y = "New positives") +
  scale_x_date(date_breaks = '5 days',
               date_labels = '%Y-%m-%d') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
Predictions using previous gam model are the following:
```{r gam predictions, echo=FALSE, message=FALSE, warning=FALSE}
gam_mod_1_pred <- predict(gam_mod1, newdata=df_extended, type="response", se.fit = TRUE)
mean(gam_mod_1_pred$se.fit)
ggplot() +
  geom_point(data = df_extended, aes(x = data, y = nuovi_positivi, color="nuovi positivi", group = 1),col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df_extended$color))])+ 
  geom_line(data = df_extended, aes(x = data, y = predict(gam_mod1, newdata=df_extended, type="response"), color = "model 4", group = 1)) +
  geom_ribbon(data = df_extended, aes(x= data, ymin=-gam_mod_1_pred$se.fit + gam_mod_1_pred$fit, ymax=gam_mod_1_pred$se.fit + gam_mod_1_pred$fit), alpha=0.5)+
  
  
  labs(x = "Time", y = "New positives") +
  scale_x_date(date_breaks = '5 days',
               date_labels = '%Y-%m-%d') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


```


# Predictions

We investigate prediction accuracy testing previous models on the next 15 days of Febraury.

### Linear model
```{r lm predictions, echo = FALSE, message=FALSE, warning=FALSE}

### lm

ggplot() +
  geom_point(data = df_extended, aes(x = data, y = nuovi_positivi, color="nuovi positivi", group = 1),col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df_extended$color))])+ 
  geom_line(data= df_extended, aes(x= data, y = predict(lm_mod5, newdata = df_extended, type="response"), color="model 5"), size=0.75)+
  labs(x = "Time", y = "New positives") +
  scale_x_date(date_breaks = '5 days',
               date_labels = '%Y-%m-%d') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_color_brewer(palette = "Dark2")
```

```{r glm predictions, echo = FALSE, message=FALSE, warning=FALSE}
ggplot() +
  geom_point(data = df_extended, aes(x = data, y = nuovi_positivi, color="nuovi positivi", group = 1),col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df_extended$color))])+ 
  geom_line(data= df_extended, aes(x= data, y = predict(glm_mod5, newdata = df_extended, type="response"), color="model 5"), size=0.75)+
  labs(x = "Time", y = "New positives") +
  scale_x_date(date_breaks = '5 days',
               date_labels = '%Y-%m-%d') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_color_brewer(palette = "Dark2")

#ggsave(paste0("images/glm_mod.png"), width = 12, height = 5, dpi = 200)
```


```{r accuracy evaluation, echo = FALSE, message=FALSE, warning=FALSE}
library("Metrics")
#lm
x1 <- 1 - smape(predict(lm_mod5,type="response", newdata= df_extended[124:137,]), df_extended$nuovi_positivi[124:137])[1] #  0.3325639

#glm poisson
x2 <- 1 - smape(predict(glm_mod5,type="response", newdata= df_extended[124:137,]), df_extended$nuovi_positivi[124:137])[1] # 0.3308621


#glm negative binomial
x3 <- 1 - smape(predict(nglmod5,type="response", newdata= df_extended[124:137,]), df_extended$nuovi_positivi[124:137])[1] # 0.3537849

#gam
x4 <- 1 - smape(predict(gam_mod1,type="response", newdata= df_extended[124:137,]), df_extended$nuovi_positivi[124:137])[1] # 0.4643648


table <- matrix(c("lm_mod5", "glm_mod5","nglmod5", "gam_mod1",x1,x2,x3,x4), ncol=2, byrow=FALSE)
colnames(table) <- c('model','smape')
table
```


# NEW COLOR

All the models previously considered overestimate the number of new positives on the first two weeks of Febraury.
This behaviour might be due to the gap between the variable color and the actual level of restrictions.
To handle this issue a new numerical variable (livello_di_restrizioni) has been added.

```{r new color, echo = FALSE, message=FALSE, warning=FALSE}

ggplot() +
  geom_point(data = df, aes(x = data, y = new_color, group = 1),col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))])+ 
  geom_line(data = df, aes(x = data, y = new_color, group = 1), size=1.6,col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df$color))]) +
  labs(x = "data", y = "livello di restrizioni") +
  scale_x_date(date_breaks = '5 days',
               date_labels = '%Y-%m-%d') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggsave(paste0( "images/restriction_level.png"), width = 16, height = 9, dpi = 150)

```

Let's see how our previous model change adding new color:

-   Linear model
```{r linear model using new color, echo = FALSE, message=FALSE, warning=FALSE}
### lm

lm_mod <- lm(nuovi_positivi ~ ricoverati_con_sintomi + nuovi_tamponi_pcr + new_color, data = df)

ggplot() +
  geom_point(data = df_extended, aes(x = data, y = nuovi_positivi, color="nuovi positivi", group = 1),col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df_extended$color))])+ 
  geom_line(data= df_extended, aes(x= data, y = predict(lm_mod, newdata = df_extended, type="response"), color="model 5"), size=0.75)+
  labs(x = "Time", y = "New positives") +
  scale_x_date(date_breaks = '5 days',
               date_labels = '%Y-%m-%d') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_color_brewer(palette = "Dark2")
```

-   Generalized linear model - Poisson
```{r glm pois using new color, echo = FALSE, message=FALSE, warning=FALSE}
### glm
glm_mod <- glm(nuovi_positivi ~ ricoverati_con_sintomi + nuovi_tamponi_pcr + new_color, data = df, family=poisson(link = "log"))

ggplot() +
  geom_point(data = df_extended, aes(x = data, y = nuovi_positivi, color="nuovi positivi", group = 1),col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df_extended$color))])+ 
  geom_line(data= df_extended, aes(x= data, y = predict(glm_mod, newdata = df_extended, type="response"), color="model 5"), size=0.75)+
  labs(x = "Time", y = "New positives") +
  scale_x_date(date_breaks = '5 days',
               date_labels = '%Y-%m-%d') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_color_brewer(palette = "Dark2")

#ggsave(paste0("images/glm_mod.png"), width = 12, height = 5, dpi = 200)
```

-   Generalized linear model - Ngeative Binomial
```{r glm nb using new color, echo = FALSE, message=FALSE, warning=FALSE}

### glm nb
nglm_mod <- glm.nb(nuovi_positivi ~ ricoverati_con_sintomi + nuovi_tamponi_pcr + new_color, data = df)

ggplot() +
  geom_point(data = df_extended, aes(x = data, y = nuovi_positivi, color="nuovi positivi", group = 1),col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df_extended$color))])+ 
  geom_line(data= df_extended, aes(x= data, y = predict(nglm_mod, newdata = df_extended, type="response"), color="model 5"), size=0.75)+
  labs(x = "Time", y = "New positives") +
  scale_x_date(date_breaks = '5 days',
               date_labels = '%Y-%m-%d') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_color_brewer(palette = "Dark2")
```

-   Generalized additive models
```{r gam using new color, echo = FALSE, message=FALSE, warning=FALSE}
### gam
gam_mod <- gam(nuovi_positivi ~ s(ricoverati_con_sintomi) + s(nuovi_tamponi_pcr) + s(new_color), data = df)

ggplot() +
  geom_point(data = df_extended, aes(x = data, y = nuovi_positivi, color="nuovi positivi", group = 1),col=c("#fc6b03","#cfcaca","#f2d729","#b3190b")[unclass(as.factor(df_extended$color))])+ 
  geom_line(data= df_extended, aes(x= data, y = predict(gam_mod, newdata = df_extended, type="response"), color="model 5"), size=0.75)+
  labs(x = "Time", y = "New positives") +
  scale_x_date(date_breaks = '5 days',
               date_labels = '%Y-%m-%d') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_color_brewer(palette = "Dark2")

```

Check for prediction accuracy:

```{r prediction accuracy evaluation, echo = FALSE, message=FALSE, warning=FALSE}
library("Metrics")
#lm
x1 <- 1 - smape(predict(lm_mod,type="response", newdata= df_extended[124:137,]), df_extended$nuovi_positivi[124:137]) #0.7969439

#glm poisson
x2 <- 1 - smape(predict(glm_mod,type="response", newdata= df_extended[124:137,]), df_extended$nuovi_positivi[124:137]) #0.8890304

#nglm
x3 <- 1 - smape(predict(nglm_mod,type="response", newdata= df_extended[124:137,]), df_extended$nuovi_positivi[124:137]) #0.8799145

#gam
x4 <- 1 - smape(predict(gam_mod,type="response", newdata= df_extended[124:137,]), df_extended$nuovi_positivi[124:137]) #0.8836843

table <- matrix(c("lm_mod", "glm_mod","nglmod", "gam_mod",x1,x2,x3,x4), ncol=2, byrow=FALSE)
colnames(table) <- c('model','smape')
table

```

Plot previous models lm and glm and highlight predictions:
```{r}

plot_predictions <- function(mod) {
  p <- ggplot() +
    geom_point(data = df_extended, aes(x = data, y = nuovi_positivi, color="nuovi positivi", group = 1),col=col_zone[unclass(as.factor(df_extended$color))])+ 
    geom_line(data = df_extended, aes(x = data, y = predict(mod, newdata = df_extended, type="response"), color = "model 1", group = 1)) + labs(x = "data", y = "nuovi positivi") +
    scale_x_date(date_breaks = '5 days', date_labels = '%Y-%m-%d') +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    scale_color_brewer(palette = "Dark2")
  
  d1<-df_extended$data[124]
  d2<-df_extended$data[137]
  
  g <- ggplot(data = df_extended)+
    geom_point( aes(x = data, y = nuovi_positivi, color="nuovi positivi", group = 1),col=col_zone[unclass(as.factor(df_extended$color))]) +
    geom_line(data = df_extended, aes(x = data, y = predict(mod, newdata = df_extended, type="response"), color = "model 1", group = 1)) + labs(x = "data", y = "nuovi positivi") +
    scale_color_brewer(palette = "Dark2") +
    scale_x_date(limits = c(d1, d2)) + scale_y_continuous(limits = c(400, 1100)) + theme_bw()+
    theme(
      legend.position = "none",
      panel.background = element_rect(fill = "transparent"), # bg of the panel
      #plot.background = element_rect(fill = "transparent", color = NA), # bg of the plot
      #panel.grid.major = element_blank(), # get rid of major grid
      #panel.grid.minor = element_blank(), # get rid of minor grid
      legend.background = element_rect(fill = "transparent"), # get rid of legend bg
      legend.box.background = element_rect(fill = "transparent"), # get rid of legend panel bg,
      axis.title.x=element_blank(),
      axis.text.x=element_blank(),
      axis.ticks.x=element_blank(),
      axis.title.y=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks.y=element_blank(),
      plot.margin = margin(t = 0,  # Top margin
                           r = 0,  # Right margin
                           b = 0,  # Bottom margin
                           l = 0) # Left margin
    )
  
  d3 <- df_extended$data[75]
  d4 <- df_extended$data[118]
  h <- p + annotation_custom(ggplotGrob(g), xmin=d3, xmax=d4, ymin=0, ymax=1000)
  return(h)
}

res <- plot_predictions(lm_mod)
res
#ggsave("images/lm_predictions.png", width = 10, height = 8, dpi = 200)

res <- plot_predictions(glm_mod)
res
#ggsave("images/glm_predictions.png", width = 10, height = 8, dpi = 200)

```


# Google data

To further improve our model we tried to introduce another covariates which accounts for the movements tracked by Google.

Linear models and generalized linear models using google data.

```{r lm glm using google data}

lm_mod_google <- lm(nuovi_positivi ~ ricoverati_con_sintomi + nuovi_tamponi_pcr + var_station + new_color, data = df) 

glm_mod_google <- glm(nuovi_positivi ~ ricoverati_con_sintomi + nuovi_tamponi_pcr + var_station + new_color, data = df, family=poisson(link = "log"))

res <- plot_predictions(lm_mod_google)
res
ggsave(paste0( "images/lm_predictions_google.png"), width = 10, height = 8, dpi = 200)


res <- plot_predictions(glm_mod_google)
res
ggsave(paste0( "images/glm_predictions_google.png"), width = 10, height = 8, dpi = 200)

#lm
x1 <- 1 - smape(predict(lm_mod_google,type="response", newdata = df_extended[124:137,]), df_extended$nuovi_positivi[124:137]) #0.7864553

#glm poisson
x2 <- 1 - smape(predict(glm_mod_google,type="response", newdata= df_extended[124:137,]), df_extended$nuovi_positivi[124:137]) 
#0.8775334

table <- matrix(c("lm_mod_google", "glm_mod_google",x1,x2), ncol=2, byrow=FALSE)
colnames(table) <- c('model','smape')
table

```


# MARS

```{r spline, echo = FALSE, message = FALSE}
library(splines)
knots <- quantile(df$ricoverati_con_sintomi)[2:4]
fit.spline <- lm (nuovi_positivi ~ bs(ricoverati_con_sintomi + new_color) + nuovi_tamponi_pcr, data = df)
summary(fit.spline)


predictions <- fit.spline %>% predict(df_extended)

ggplot() +
  geom_point(data = df_extended, aes(data, nuovi_positivi))+
  geom_line(data = df_extended, aes(x= data, y = predictions, color="Spline model"), size=0.75)+
  scale_color_brewer(palette = "Dark2")+
  labs(x = "Time", y = "New positives") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Compare fitting using glm and fitting using mars with respect to ricoverati_con_sintomi covariate.

```{r mars, echo = FALSE, message = FALSE}
library(earth)
earth.mod <- earth(nuovi_positivi ~ ricoverati_con_sintomi + nuovi_tamponi_pcr + new_color, data = df, degree = 5, nk = 8)

plotmo(earth.mod)
plot(earth.mod)
summary(earth.mod, digits = 2, style = "pmax")

ggplot() +
  geom_point(data = df, aes(x = ricoverati_con_sintomi, y = nuovi_positivi, color="nuovi positivi", group = 1),col=col_zone[unclass(as.factor(df$color))])+ 
  geom_line(data = df, aes(x= ricoverati_con_sintomi, y = predict(earth.mod, newdata = df, type="response"), color="model 5"), size=0.75)+
  labs(x = "Time", y = "New positives") +

  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_color_brewer(palette = "Dark2")

ggsave("images/mars_ricoverati_con_sintomi.png", width = 16, height = 9, dpi = 200)

ggplot() +
  geom_point(data = df_extended, aes(x = ricoverati_con_sintomi, y = nuovi_positivi, color="nuovi positivi", group = 1),col=col_zone[unclass(as.factor(df_extended$color))])+ 
  geom_line(data= df_extended, aes(x= ricoverati_con_sintomi, y = predict(glm_mod, newdata = df_extended, type="response"), color="model 5"), size=0.75)+
  labs(x = "ricoverati con sintomi", y = "nuovi positivi") +

  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_color_brewer(palette = "Dark2")

ggsave("images/glm_ricoverati_con_sintomi.png", width = 16, height = 9, dpi = 200)

```
Predictions using mars.

```{r mars predictions, echo = FALSE, message = FALSE}


1 - smape(predict(earth.mod,type="response", newdata= df_extended[124:137,]), df_extended$nuovi_positivi[124:137]) 

res <- plot_predictions(earth.mod)
res
ggsave("images/mars_predictions.png", width = 16, height = 9, dpi = 200)
```